GEE models

### 2017-07-20 doc/esm-methods.Rmd
minor clarifications to pls in esm methods
clarified dag and qic in esm methods

For the confounders, they were chosen based on previous literature, from 
directed acyclic graph (DAG) recommendations, and from QIC. The DAG
recommendations were obtained from using the DAGitty software,
http://dagitty.net/. The final GEE model selected as best fitting was negligibly
different when using insulin sensitivity or beta-cell function as an outcome and
as such the same covariates were chosen for models with each of the outcomes.

Partial least squares (PLS) regression is a technique similar to principal
component analysis (PCA) that is designed to extract meaningful information from
multivariate, high dimensionality data (e.g. as in metabolomic and other -omic type 
analyses). These types of methods try to extract as much of the variance in the 
data into a smaller number of components or factors. The difference between PCA 
and PLS is that PLS is a supervised method, while PCA is not. This means that
PLS uses a response variable(s), i.e. an outcome or **Y**, to describe the
variation in the predictors (**X**) while PCA only describes the variation 
inherent in **X**. Because of this, PLS can be better at predicting the 
contributions of predictor variables against an outcome variable.

>>>>

For the covariates, they were chosen based on previous literature, from 
directed acyclic graph (DAG) recommendations, and from QIC. The DAG
recommendations were obtained from using the DAGitty software,
http://dagitty.net/. DAG is a technique that allows for identifying potential
confounding causal pathways and to limit potential bias introduced from
adjusting for colliding variables. A collider variable is a variable that is
determined from two or more other variables, such that in causal graph modeling,
the arrows from the other variables appear to 'collide' into the variable. QIC 
is a technique that balances the goodness of fit of the model with its relative
complexity (i.e. number of covariates) against other models being compared. This
identifies a specific model from a number of models that is the 'best' fit and
least complex of all the models. QIC and other information criterion methods are
used in conjunction with other techniques such as literature and DAG based. The 
final GEE model selected as best fitting was negligibly different when using 
insulin sensitivity or beta-cell function as an outcome and as such the same 
covariates were chosen for models with each of the outcomes.

Partial least squares (PLS) regression is a technique similar to principal
component analysis (PCA) that is designed to extract meaningful information from
multivariate, high dimensionality data (e.g. as in metabolomic and other -omic type 
analyses). These types of methods try to extract as much of the variance in the 
data into a smaller number of components or factors. The difference between PCA 
and PLS is that PLS is a supervised method, while PCA is not. This means that
PLS uses a response variable(s), i.e. an outcome or **Y**, to 'supervise' (or
constrain) the variation inherent in the predictors (**X**) while PCA only
describes the variation inherent in **X**. Because of this, PLS can be better at
predicting the contributions of predictor variables against an outcome variable.
